---
title: "Clean WeatherUnderground of NC on JJA 2021"
format: html
editor: visual
---

The following code generates clean CWS data on the Triangle.

#### Load brassens library

```{r}
devtools::load_all()
```

### Open NC data

```{r}
nc_shp <- paste0(
  "../input/NC_county_boundary/",
  "North_Carolina_State_and_County_Boundary_Polygons.shp"
) |>
  terra::vect()
nc_inv <- read.csv("../input/weatherunderground/Summer2021_inventory.csv") |>
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326, remove = FALSE) |>
  sf::st_transform(terra::crs(nc_shp))

nc_inv$ts_utc <- as.POSIXct(nc_inv$ts_utc,
                            format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
nc_inv$te_utc <- as.POSIXct(nc_inv$te_utc,
                            format = "%Y-%m-%d %H:%M:%S", tz = "UTC")

imp_path <- paste0("../input/US_2019_imperviousness/",
                   "nlcd_2019_impervious_l48_20210604.img")
imp_us <- terra::rast(imp_path) |>
  terra::aggregate(fact = 20, fun = "mean", na.rm = TRUE) |>
  terra::project("epsg:4326")
nc_shp_proj <- terra::project(nc_shp, "epsg:4326")
imp_nc <- terra::crop(imp_us,
                      nc_shp_proj,
                      overwrite = TRUE,
                      mask = TRUE,
                      snap = "out")
```

### Create configuration list for the pipeline

```{r}
config <- list(
  ts = as.POSIXct("2021-06-01 00:00:00", tz = "UTC"),
  te = as.POSIXct("2021-08-31 23:59:59", tz = "UTC"),
  area = nc_shp,
  wu_inv = nc_inv
)
```

### Open GHCNh data

```{r, eval = FALSE}
setwd("../")
start_time <- Sys.time()
ghcnh <- download_ghcnh(config$ts, config$te, config$area)
saveRDS(
  ghcnh,
  paste0(
    "./output/case_study_nc_jja2021/ghcnh_nc_",
    lubridate::date(config$ts),
    "_",
    lubridate::date(config$te),
    ".rds"
  )
)
end_time <- Sys.time()
end_time - start_time
cat(
  "There are ",
  length(unique(ghcnh$site_id)),
  " GHCNh stations in NC for these dates."
)
```

```{r}
ghcnh <- readRDS(paste0(
  "../output/case_study_nc_jja2021/ghcnh_nc_",
  lubridate::date(config$ts),
  "_",
  lubridate::date(config$te),
  ".rds"
))
```

### Pipeline

Apply the pipeline by batches of 2 weeks

```{r}
setwd("../")
start_time <- Sys.time()
batches <- seq(config$ts, config$te, by = "2 weeks")
for (i in 1:(length(batches) - 1)) {
  wu_list_cc <- NULL
  ts <- batches[i]
  te <- batches[i + 1]
  cat(
    "Process cleaning for ",
    strftime(ts, format = "%Y-%m-%d %H:%M:%S", tz = "UTC"),
    " - ",
    strftime(te, format = "%Y-%m-%d %H:%M:%S", tz = "UTC"),
    "\n"
  )
  # run brassens pipeline on WeatherUnderground between ts and te
  ghcnh_week <- ghcnh[which(dplyr::between(ghcnh$time, ts, te)), ]
  wu_list_cc <- load_wu(ts, te, config$area, config$wu_inv) |>
    format_wu() |>
    clean_cws_large(config$area, epsg_m = "epsg:32119") |>
    calib_cws(ref = ghcnh_week, max_dist = 10000)
  # save result
  saveRDS(wu_list_cc, paste0(
    "./output/case_study_nc_jja2021/wu_nc_cleaned_and_calibrated_",
    lubridate::date(ts),
    "_",
    lubridate::date(te),
    ".rds"
  ))
}
end_time <- Sys.time()
end_time - start_time
```

Add last (incomplete) batch from 08/24 to 08/31

```{r}
setwd("../")
wu_list_cc <- NULL
ts <- as.POSIXct("2021-08-24 00:00:00", tz = "UTC")
te <- as.POSIXct("2021-08-31 23:59:59", tz = "UTC")
cat(
  "Process cleaning for ",
  strftime(ts, format = "%Y-%m-%d %H:%M:%S", tz = "UTC"),
  " - ",
  strftime(te, format = "%Y-%m-%d %H:%M:%S", tz = "UTC"),
  "\n"
)
# run brassens pipeline on WeatherUnderground between ts and te
ghcnh_week <- ghcnh[which(dplyr::between(ghcnh$time, ts, te)), ]
wu_list_cc <- load_wu(ts, te, config$area, config$wu_inv) |>
  format_wu() |>
  clean_cws_large(config$area, epsg_m = "epsg:32119") |>
  calib_cws(ref = ghcnh_week, max_dist = 10000)
# save result
saveRDS(wu_list_cc, paste0(
  "./output/case_study_nc_jja2021/wu_nc_cleaned_and_calibrated_",
  lubridate::date(ts),
  "_",
  lubridate::date(te),
  ".rds"
))
```

### Map squares on NC

Squares of 100km\*100km to apply CrowdQC+

```{r}
nc <- nc_shp |> terra::project(y = "epsg:32119") # linear unit of nc_shp in m
r <- terra::rast(nc, res = 100000)
terra::values(r) <- 1:terra::ncell(r)
v <- terra::mask(r, nc) |>
  terra::as.polygons()
```

Plot polygons with imperviousness

```{r}
imp_nc_proj <- terra::project(x = imp_nc, y = "epsg:32119")

ggplot2::ggplot() +
  tidyterra::geom_spatraster(data = imp_nc_proj) +
  tidyterra::geom_spatvector(data = nc,
                             color = "white",
                             linewidth = .5,
                             fill = NA) +
  tidyterra::geom_spatvector(data = v,
                             color = "red",
                             linewidth = .5,
                             fill = NA) +
  scale_fill_gradientn(
    colours = c("black",
                "floralwhite",
                "lightpink",
                "deeppink1",
                "darkorchid1"),
    na.value = NA,
    breaks = seq(0, 100, 10),
    limits = c(0, 100)
  ) +
  labs(fill = "imperviousness (%)") +
  ggspatial::annotation_scale(
    location = "bl",
    pad_x = grid::unit(1, "cm"),
    pad_y = grid::unit(1, "cm"),
    height = grid::unit(0.30, "cm"),
    text_cex = 1
  ) +
  ggspatial::annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = grid::unit(0.2, "cm"),
    pad_y = grid::unit(0.2, "cm")
  ) +
  ggplot2::theme(
    axis.text = ggplot2::element_text(size = 12),
    plot.caption = ggplot2::element_text(size = 10),
    legend.text = ggplot2::element_text(size = 12),
    legend.title = ggplot2::element_text(size = 12),
    panel.background = ggplot2::element_rect(fill = "white"),
    panel.grid.major = ggplot2::element_line(colour = "grey")
  )
ggsave("../graphs/map_nc_cut_80kmx80km_squares.png",
  width = 7,
  height = 3
)
```

Open one period

```{r}
f <- paste0("../output/case_study_nc_jja2021/",
            "wu_nc_cleaned_and_calibrated_2021-06-01_2021-06-15.rds")
wu_list_cc <- readRDS(f)
wu_cc <- wu_list_cc$obs
```

Maps for the entire period

```{r}
ts <- range(wu_cc$time)[1]
te <- range(wu_cc$time)[2]
period <- seq(ts, te, by = "1 hour")
period <- lubridate::with_tz(period,
  "America/New_York",
  format = "%Y-%m-%d %H:%M:%S"
)
mint <- floor(min(wu_cc$temp_cal, na.rm = TRUE))
maxt <- ceiling(max(wu_cc$temp_cal, na.rm = TRUE))
for (p in period) {
  cat(strftime(p, format = "%Y-%m-%d %H:%M:%S"), "\n")

  plot <- ggplot() +
    tidyterra::geom_spatraster(data = imp_nc) +
    scale_fill_gradientn(colours = c("white", "grey"), na.value = NA) +
    geom_sf(
      data = wu_cc[which(wu_cc$time == p), ],
      aes(geometry = geometry, color = temp_cal)
    ) +
    ggtitle(strftime(p, format = "%Y-%m-%d %H:%M:%S")) +
    labs(fill = "imp (%)", color = "T (°C)") +
    tidyterra::scale_color_whitebox_c(
      palette = "bl_yl_rd",
      n.breaks = 12,
      limits = c(mint, maxt),
      guide = guide_legend(reverse = TRUE)
    )
  ggsave(
    plot = plot,
    filename = paste0(
      "../output/case_study_nc_jja2021/graphs/pmap_",
      strftime(p, format = "%Y-%m-%d %H:%M:%S"),
      ".png"
    ),
    width = 12,
    height = 5,
    dpi = 300
  )
  plot
}
```

Save cleaned and calibrated cws data in .rds file for mercury analysis

```{r}
saveRDS(wu_list_cc, file = paste0(
  "../output/wu_nc_cleaned_calibrated_",
  lubridate::date(config$ts),
  "_",
  lubridate::date(config$te),
  ".rds"
))
```
